|

### 2023-04-16 appendix.Rmd
added section to appendix exploring the rejection rate by sig level

```{r}
upper_lower_p %>%
  left_join(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value = first(p_value))) %>%
  group_by(modified) %>%
  count(adjusted_reject = adjusted_p_value <= 0.05, reject = p_value <= 0.05) %>%
  complete(adjusted_reject, reject, fill = list(n = 0)) %>%
  ungroup() %>%
  filter(adjusted_reject != reject) %>%
  mutate(case = ifelse(!adjusted_reject, "reject to not reject", "not reject to reject")) %>%
  rename(multiplier = modified) %>%
  select(-adjusted_reject, -reject) %>%
  pivot_wider(names_from = case, values_from = n) %>%
  mutate(reject_not_reject_per = round(`reject to not reject`/279*100, 2)) %>%
  mutate(not_reject_reject_per = round(`not reject to reject`/279*100, 2)) %>%
  select(multiplier,
         `reject to not reject`,
         reject_not_reject_per,
         `not reject to reject`,
         not_reject_reject_per) %>%
  kableExtra::kable(format = "latex", 
                    booktabs = TRUE,
                    col.names = c("Multiplier",
                                  "Reject to not reject",
                                  "%",
                                  "Not reject to reject",
                                  "%"),
                    label = "sensitivity", 
                    caption = "Examining how decisions might change if $\\hat{\\alpha}$ was different. Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is multiplied by a multiplier.") %>%
  kableExtra::kable_styling(latex_options = "scale_down")
```

@@

```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution. The power will incrase as the number of evaluations increases, but the margin will decrase.", fig.width=6, fig.height=4, out.width="70%"}
eval_n <- function(n_eval, boot) {
  map_df(1:boot, function(boot_id) {
  vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  filter(type == "polynomial") %>%
  group_by(unique_lineup_id) %>%
  slice_sample(n = n_eval) %>%
  ungroup() %>%
  calc_p_value_multi(lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha) %>%
  mutate(boot_id = boot_id)
}) %>%
  left_join(select(vi_survey, unique_lineup_id, effect_size)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(-unique_lineup_id, -p_value) %>%
  mutate(n_eval = n_eval)
}

@@

\newpage
# Experiment setup

>>>>

```{r}
upper_lower_p %>%
  left_join(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value = first(p_value))) %>%
  group_by(modified) %>%
  count(adjusted_reject = adjusted_p_value <= 0.05, reject = p_value <= 0.05) %>%
  complete(adjusted_reject, reject, fill = list(n = 0)) %>%
  ungroup() %>%
  filter(adjusted_reject != reject) %>%
  mutate(case = ifelse(!adjusted_reject, "reject to not reject", "not reject to reject")) %>%
  rename(multiplier = modified) %>%
  select(-adjusted_reject, -reject) %>%
  pivot_wider(names_from = case, values_from = n) %>%
  mutate(reject_not_reject_per = round(`reject to not reject`/279*100, 2)) %>%
  mutate(not_reject_reject_per = round(`not reject to reject`/279*100, 2)) %>%
  select(multiplier,
         `reject to not reject`,
         reject_not_reject_per,
         `not reject to reject`,
         not_reject_reject_per) %>%
  kableExtra::kable(format = "latex", 
                    booktabs = TRUE,
                    col.names = c("Multiplier",
                                  "Reject to not reject",
                                  "%",
                                  "Not reject to reject",
                                  "%"),
                    label = "sensitivity", 
                    caption = "Examining how decisions might change if $\\hat{\\alpha}$ was different. Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is multiplied by a multiplier.",
                    linesep = c("", "", "\\addlinespace"))
```

@@

```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution. The power will increase as the number of evaluations increases, but the margin will decrease.", fig.width=6, fig.height=4, out.width="70%"}
eval_n <- function(n_eval, boot) {
  map_df(1:boot, function(boot_id) {
  vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  filter(type == "polynomial") %>%
  group_by(unique_lineup_id) %>%
  slice_sample(n = n_eval) %>%
  ungroup() %>%
  calc_p_value_multi(lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha) %>%
  mutate(boot_id = boot_id)
}) %>%
  left_join(select(vi_survey, unique_lineup_id, effect_size)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(-unique_lineup_id, -p_value) %>%
  mutate(n_eval = n_eval)
}

@@

## Conventional test rejection rate for varying significance levels

In the main paper, Sections 5.1 and 5.2 compared the power, and the decisions made by the conventional tests and the visual test. The power curves for the visual test is effectively a right-shift from the conventional test. The effect is that the visual test rejects less often than the conventional test, at the same significance level. We also saw that the visual test rejected a subset of those that the conventional tests rejected. This means that they agreed quite well - only residual plots rejected by the contentional tests were rejected by the visual test. There was little disagreement, where residual plots not rejected by the conventional test were rejected by the visual test. The question arises whether the decisions made conventional test could be made similar to that of the visual test by reducing the significance level. Reducing the significance level from 0.05, to 0.01, 0.001, ... will have the effect of rejecting fewer of the residual plots.

It would be interesting if a different conventional test significance level results in both the visual tests and conventional tests reject only the same residual plots, and fails to reject the same residual plots. This would be a state where both systems agree perfectly. Figure \ref{fig:rej-rates} examines this. Plot A shows the percentage of residual plots rejected by the visual test, given the conventional test rejected (solid lines) or failed to reject (dashed lines). The vertical grey line marks significance level 0.05. When the signicance level gets smaller, the it is possible te see that the visual tests reject (nearly) 100% of the time that the conventional test rejects. However, there is not agreement, because the visual tests also increasingly reject residual plots where the conventional test failed to reject. Plot B is comparable to an ROC curve, where the percentage visual test rejection conditional on conventional test decision is plotted: Reject conditional on reject is plotted against reject conditional on fail to reject, for different significance levels. The non-linearity pattern results are close to being ideal, that the percentage of reject relative to fail to reject increases very slowly as the reject relative to reject converges to 100. The heteroskedasticity pattern is more problematic, and shows that the cost of rejecting less with the conventional test is disagreement with the visual test.

```{r rej-rates, fig.pos="t!", fig.width=8, fig.height=5, out.width="100%", fig.cap="Changing the significance level of the conventional test will change the rejection rate. A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed). As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100. The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too. B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test. Many of the residual plots not rejected by the conventional test are rejected by the visual test. It isn't possible to vary the significance level of the conventional test to match the decisions made by the visual test."}
# compute rejection rate for different significance levels, holding visual test constant
library(ggpubr)
nl_rates <- NULL
h_rates <- NULL
for (i in c(0.1, 0.05, 0.01, 0.001, 0.0001, 0.00001, 0.000001, 0.0000001, 0.00000001, 0.000000001,
            0.0000000001, 0.00000000001)) {
p_value_cmp_dat <- vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(conv_reject = ifelse(conventional_p_value <= i, "Reject", "Not"),
         reject = ifelse(p_value <= 0.05, "Reject", "Not")) %>%
  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
         reject = factor(reject, levels = c("Reject", "Not"))) %>%
  select(type, conv_reject, reject) 
nl <- p_value_cmp_dat %>% 
  filter(type == "non-linearity") %>% 
  count(conv_reject, reject, .drop=FALSE) %>%
  summarise(rej_pct=round(n[1]/(n[1]+n[2])*100, 0),
            nrej_pct=round(n[3]/(n[3]+n[4])*100, 0)) %>%
  bind_cols(alpha=i)
  nl_rates <- bind_rows(nl_rates, nl)
  # Heteroskedasticity 
  h <- p_value_cmp_dat %>% 
  filter(type == "heteroskedasticity") %>% 
  count(conv_reject, reject, .drop=FALSE) %>%
  summarise(rej_pct=round(n[1]/(n[1]+n[2])*100, 0),
            nrej_pct=round(n[3]/(n[3]+n[4])*100, 0)) %>%
  bind_cols(alpha=i)
  h_rates <- bind_rows(h_rates, h)
}
nl_rates <- nl_rates %>% mutate(type = "non-linearity")
h_rates <- h_rates %>% mutate(type = "heteroskedasticity")
rates <- bind_rows(nl_rates, h_rates) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity")))
rej_by_sig <- ggplot(rates) +
  geom_vline(xintercept=0.05, colour="grey50") +
  geom_line(aes(x=alpha, y=rej_pct, colour=type), alpha=0.8) +
  geom_line(aes(x=alpha, y=nrej_pct, colour=type), linetype=2) +
  scale_x_log10(breaks=c(0.1, 0.001, 0.00001, 0.0000001, 0.000000001, 0.00000000001)) + 
  ylab("Visual test rejects (%)") + xlab("Significance level") +
  scale_colour_brewer("", palette = "Dark2") +
  ggtitle("A")
roc <- ggplot(rates) +
  geom_path(aes(x=nrej_pct, y=rej_pct, colour=type)) +
  ylab("Reject | Reject (%)") + xlab("Reject | Fail to reject (%)") +
  ylim(0, 100) +
  scale_colour_brewer("", palette = "Dark2") +
  coord_equal() +
  ggtitle("B")
ggarrange(rej_by_sig, roc, ncol=2, common.legend=TRUE, legend = "bottom")
```

# Experiment setup

### 2023-04-16 appendix.tex
added section to appendix exploring the rejection rate by sig level

\caption{\label{tab:sensitivity}Examining how decisions might change if $\hat{\alpha}$ was different. Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\hat{\alpha}$ is multiplied by a multiplier.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rrrrr}
\toprule
Multiplier & Reject to not reject & \% & Not reject to reject & \%\\
\midrule
0.500 & 7 & 2.51 & 0 & 0.00\\
0.750 & 4 & 1.43 & 0 & 0.00\\
0.875 & 3 & 1.08 & 0 & 0.00\\
1.125 & 0 & 0.00 & 3 & 1.08\\
1.250 & 0 & 0.00 & 4 & 1.43\\
\addlinespace
1.500 & 0 & 0.00 & 5 & 1.79\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution. The power will incrase as the number of evaluations increases, but the margin will decrase.}\label{fig:n_eval}
\end{figure}

@@

>>>>

\caption{\label{tab:sensitivity}Examining how decisions might change if $\hat{\alpha}$ was different. Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\hat{\alpha}$ is multiplied by a multiplier.}
\centering
\begin{tabular}[t]{rrrrr}
\toprule
Multiplier & Reject to not reject & \% & Not reject to reject & \%\\
\midrule
0.500 & 7 & 2.51 & 0 & 0.00\\
0.750 & 4 & 1.43 & 0 & 0.00\\
0.875 & 3 & 1.08 & 0 & 0.00\\
\addlinespace
1.125 & 0 & 0.00 & 3 & 1.08\\
1.250 & 0 & 0.00 & 4 & 1.43\\
1.500 & 0 & 0.00 & 5 & 1.79\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution. The power will increase as the number of evaluations increases, but the margin will decrease.}\label{fig:n_eval}
\end{figure}

@@

\hypertarget{conventional-test-rejection-rate-for-varying-significance-levels}{%
\subsection{Conventional test rejection rate for varying significance
levels}\label{conventional-test-rejection-rate-for-varying-significance-levels}}

In the main paper, Sections 5.1 and 5.2 compared the power, and the
decisions made by the conventional tests and the visual test. The power
curves for the visual test is effectively a right-shift from the
conventional test. The effect is that the visual test rejects less often
than the conventional test, at the same significance level. We also saw
that the visual test rejected a subset of those that the conventional
tests rejected. This means that they agreed quite well - only residual
plots rejected by the contentional tests were rejected by the visual
test. There was little disagreement, where residual plots not rejected
by the conventional test were rejected by the visual test. The question
arises whether the decisions made conventional test could be made
similar to that of the visual test by reducing the significance level.
Reducing the significance level from 0.05, to 0.01, 0.001, \ldots{} will
have the effect of rejecting fewer of the residual plots.

It would be interesting if a different conventional test significance
level results in both the visual tests and conventional tests reject
only the same residual plots, and fails to reject the same residual
plots. This would be a state where both systems agree perfectly. Figure
\ref{fig:rej-rates} examines this. Plot A shows the percentage of
residual plots rejected by the visual test, given the conventional test
rejected (solid lines) or failed to reject (dashed lines). The vertical
grey line marks significance level 0.05. When the signicance level gets
smaller, the it is possible te see that the visual tests reject (nearly)
100\% of the time that the conventional test rejects. However, there is
not agreement, because the visual tests also increasingly reject
residual plots where the conventional test failed to reject. Plot B is
comparable to an ROC curve, where the percentage visual test rejection
conditional on conventional test decision is plotted: Reject conditional
on reject is plotted against reject conditional on fail to reject, for
different significance levels. The non-linearity pattern results are
close to being ideal, that the percentage of reject relative to fail to
reject increases very slowly as the reject relative to reject converges
to 100. The heteroskedasticity pattern is more problematic, and shows
that the cost of rejecting less with the conventional test is
disagreement with the visual test.

\begin{figure}[t!]

{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/rej-rates-1}

\caption{Changing the significance level of the conventional test will change the rejection rate. A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed). As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100. The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too. B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test. Many of the residual plots not rejected by the conventional test are rejected by the visual test. It isn't possible to vary the significance level of the conventional test to match the decisions made by the visual test.}\label{fig:rej-rates}
\end{figure}

### 2023-04-16 paper.Rmd
added section to appendix exploring the rejection rate by sig level

Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals, for non-linearity and heteroskedasticity, respectively. In each case the power of visual test is re-calculated for multiple bootstrap samples leading to the many (orange) curves. The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale. To facilitate visual calibration of effect size values with the corresponding diagnostic plots, a sequence of example residual plots with increasing effect sizes is provided at the bottom of these figures. These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots. The horizontal lines of dots at 0 and 1 represent the non-reject or reject made by visual tests for each lineup.

@@

Figures \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals. The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity. The power curves for the visual test are effectively a shift right from that of the RESET test. This means that the RESET test will reject a lower effect size (less structure) than the visual test, but otherwise the performance will be similar. In other words, the RESET test is more sensitive than the visual test. This is not necessarily a good feature, for the purposes of diagnosing model defects. If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25. As expected the BP and SW tests have much lower power - they are not designed to detect non-linearity.

>>>>

Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals, for non-linearity and heteroskedasticity, respectively. In each case the power of visual test is calculated for multiple bootstrap samples leading to the many (orange) curves. The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale. To facilitate visual calibration of effect size values with the corresponding diagnostic plots, a sequence of example residual plots with increasing effect sizes is provided at the bottom of these figures. These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots. The horizontal lines of dots at 0 and 1 represent the non-reject or reject made by visual tests for each lineup.

@@

Figures \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals. The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity. (Note that the BP and SW tests have much lower power, which is expected because they are not designed to detect non-linearity.) The bootstrapped power curves for the visual test are effectively a shift right from that of the RESET test. This means that the RESET test will reject at a lower effect size (less structure) than the visual test, but otherwise the performance will be similar. In other words, the RESET test is more sensitive than the visual test. This is not necessarily a good feature for the purposes of diagnosing model defects. If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25, for a significance level of 0.05. It is also interesting to note that setting the significance level much lower, .001, .0001, .00001, ... for the RESET test, so that it rejects less does shift the power curve towards that of the visual test. However, it never reaches the matches the performance of the visual test.
